Generating the Blueprints of the Java Ecosystem

### 2015-02-27 msr.tex
updated abstract; intro; dataset construction

We present a dataset that contains popular metrics, calculated from the analysis of a large collection of software artifacts written in Java. All artifacts were taken from the {\it Maven Central Repository}~\cite{MAVEN}. Maven is a build automation tool used primarily for Java projects, and it is maintained by the Apache Software Foundation~\cite{MAVEN}. To describe the software project being built, its dependencies, and the build order, Maven uses {\sc xml}. The central repository contains more than 400,000 {\sc jar}s, in a variety of programming languages such as Java, Clojure, Groovy, and Scala. All supported languages use the {\sc jvm} platform as their runtime environment. To build a software component, it dynamically downloads Java libraries and other plug-ins from the Maven central repository, and stores them in a local cache. The repository can be updated with new projects and also with new versions of existing projects that can depend on other versions.

@@

In this paper we present: a) the construction process to obtain the collection of the metrics results that the three aforementioned tools produced for {\bf 22730} {\sc jar}s, b) our dataset and c) how researchers and practitioners can use the dataset and produce meaningful results.

@@

The dataset construction process is illustrated in Figure~\ref{fig:dataset-construction}. Initially, a snapshot of the Maven repository was downloaded locally. The repository contains various projects that contain several versions. In our experiment, we only used Java projects. For every Java project
%gousiosg: say how you identified java projects. Previous datasets failed to
% do this, therefore it is a strong point
all versions were filtered out and only the latest was kept.

Every project consists of two {\sc jar} files, one that contains a compiled version of the project and one that contains the source code. The projects that did not had both binary and source jars, were excluded from the experiment.

When the selection process was finished, we created a series of processing tasks based on the containing all the selected {\sc jar}s, and added them to a task queue mechanism.  Then we executed a number of workers
% dimitro@bkarak: written in?
% gousiosg: ref dimitro's paper and say you followed the same process
that checked out tasks from the queue, applied the static analysis tools on each {\sc jar} and stored the results to the data repository (a {\sc m}y{\sc sql} database system). The tools that were applied on every {\sc jar} were the following.

\begin{itemize}
  \item \textbf{CKJM-ext} \textit{(ckjm)}~\cite{Spi05g}
%is an open source tool that was developed by Diomidis Spinellis and then extended by Marian Jureczko. It
calculates many software complexity metrics, including the \textit{Chidamber and Kemmerer} set of object-oriented metrics \cite{CHKE94}. The version of the tool that we used for this experiment can be found in GitHub~\cite{CKJM}.

@@

% When the selection process was finished, the {\sc jar}s were downloaded from the central repository and with the help of a queue, delivered into a series of workers, which executed the tools and provided a series of measurements stored in {\sc json} form. The result files were validated by scripts and then the data were imported in a MySQL database.

@@

Table \ref{tbl:oss-size-metrics} presents several size metrics regarding the final corpus we used. It included 11,365 projects, with more than 110 million lines of code. Overall the dataset contains almost 33 million unique measurements.

@@

The methodology to identify and calculate the {\sc dsl} metrics was the following: A set of standard {\sc dsl} application libraries was identified, and the source was scanned for specific \textit{import} statements e.g. \textit{java.util.regex}, which indicated that the standard package that implements regular expressions was used, thus regular expressions were used. If a package is detected in the source code, then the project will be tagged that it uses one (1) {\sc dsl}. Consequently, if a project has {\sc dsl} count four (4), then it means that four different application libraries were detected during the source code scan. Table \ref{tbl:dsl-list} lists the selected {\sc dsl}s application libraries. Note that we focused on libraries that were  included as part of the Java {\sc sdk}.

@@

In Section~\ref{sec:dsl}, we describe a small experiment based on these metrics that measures the popularity of {\sc dsl} usage for the dataset and therefore for the Java ecosystem in general.

>>>>

We present a dataset that contains popular metrics, calculated from the analysis of a large collection of software artifacts written in Java. All artifacts were taken from the {\it Maven Central Repository}~\cite{MAVEN}. Maven is a build automation tool used primarily for Java projects, and it is maintained by the Apache Software Foundation~\cite{MAVEN}. To describe the software project being built, its dependencies, and the build order, Maven uses {\sc xml}. The central repository contains more than 400,000 {\sc jar}s, in a variety of programming languages such as Java, Clojure, Groovy, and Scala. All supported languages use the {\sc jvm} platform as their runtime environment. To build a software component, first by using the maven accompanying tools, dynamically downloads Java libraries and other plug-ins from the Maven central repository, and stores them in a local cache. The repository can be updated with new projects and also with new versions of existing projects that can depend on other versions.

@@

In this paper we present: a) the construction process to obtain the collection of the metrics results that the three aforementioned tools produced for 22730 {\sc jar}s, b) our dataset and c) how researchers and practitioners can use the dataset and produce meaningful results.

@@

The dataset construction process is illustrated in Figure~\ref{fig:dataset-construction}. Initially, a snapshot of the Maven repository was downloaded locally. The repository contains various projects that contain several versions. In our experiment, we only used Java projects. For every Java project
all versions were filtered out and only the latest was kept. The process to identify them was the following; Every project consists of two {\sc jar} files, one that contains a compiled version of the project and one that contains the source code. The projects that did not had both binary and source jars, were excluded from the experiment. When the source {\sc jar} file was downloaded, it was scanned for Java source files. If source files were present in the archive, then the project was flagged as Java project and considered valid.

When the selection process was finished, we created a series of processing tasks based on the containing all the selected {\sc jar}s, and added them to a task queue mechanism. Then we executed a number of workers that checked out tasks from the queue, applied the static analysis tools on each {\sc jar} and stored the results to the data repository (a {\sc m}y{\sc sql} database system). The tools that were applied on every {\sc jar} were the following. The workers were written in Python. In addition, there were several shell scripts (bash) that performed several checks and sanitisation tasks.

\begin{itemize}
  \item \textbf{CKJM-ext} \textit{(ckjm)}~\cite{Spi05g} that calculates many software metrics, including the \textit{Chidamber and Kemmerer} set of object-oriented metrics \cite{CHKE94}. The version of the tool that we used for this experiment can be found in GitHub~\cite{CKJM}.

@@

@@

Table \ref{tbl:oss-size-metrics} presents several size metrics for our dataset. It includes 11,365 projects, with more than 74 million lines of code and almost 33 million unique measurements.

@@

The methodology to identify and calculate the {\sc dsl} metrics was the following: A set of standard {\sc dsl} application libraries was identified, and the source was scanned for specific \textit{import} statements e.g. \textit{java.util.regex}, which indicated that the standard package that implements regular expressions was used, thus regular expressions were used in the project. Table \ref{tbl:dsl-list} lists the selected {\sc dsl}s application libraries. Note that we focused on libraries that were  included as part of the Java {\sc sdk}.

@@

In Section~\ref{sec:dsl}, we describe an experiment based on these metrics that measures the popularity of {\sc dsl} usage for the dataset and therefore for the Java ecosystem in general.
